---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


# Introduction
In this lecture, we'll cover some fundamental concepts in R programming, including lists, functions, and libraries.

## Lists
A list is a versatile data structure in R that can contain elements of different types.

```{r}
# Creating a list
my_list <- list(name = "John", age = 25, is_student = TRUE)
my_list
```

Practice Questions:

Create a list named fruit_list containing three fruits of your choice.
Access the second element of the fruit_list.

# Creating a simple function
```{r}
add_numbers <- function(x, y) {
  result <- x + y
  return(result)
}

# Using the function
result_sum <- add_numbers(10, 5)
result_sum
```

Practice Questions:

Write a function named multiply_numbers that multiplies two given numbers.
Use the multiply_numbers function to find the product of 7 and 3.

# Installing and loading a library
```{r}
install.packages("ggplot2")
library(ggplot2)

# Using a function from the library
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point()

```


Install and load the dplyr library.
Use the head function from the dplyr library to display the first 5 rows of the mtcars dataset.
